\chapter{UX - UI}

View online \href{http://magizbox.com/training/computer_science/site/ux/}{http://magizbox.com/training/computer_science/site/ux/}

1. Design Principles
UI Design Do’s and Don’ts Android Design Principles

2. Design Trends
2.1 Material Design 1
components

We challenged ourselves to create a visual language for our users that synthesizes the classic principles of good design with the innovation and possibility of technology and science. This is material design. This spec is a living document that will be updated as we continue to develop the tenets and specifics of material design.



Tools

materialpalette.com
Icon: fa2png
UI Components


Data Binding
Transclusion
Directive - Fragments

Messaging
Intent Android 1

Intents are asynchronous messages which allow application components to request functionality from other Android components. Intents allow you to interact with components from the same applications as well as with components contributed by other applications. For example, an activity can start an external activity for taking a picture.

Intents are objects of the android.content.Intent type. Your code can send them to the Android system defining the components you are targeting. For example, via the startActivity() method you can define that the intent should be used to start an activity.

An intent can contain data via a Bundle. This data can be used by the receiving component.

Style & Theme
Android Development: Explaining Styles and Themes, https://m.youtube.com/#/watch?v=MXpG3Xq0w3A

Responsive Design
Support Multi Screen 2

Intent Android ↩

Support Multi Screen ↩


