\chapter{Scala}

View online \href{http://magizbox.com/training/scala/site/}{http://magizbox.com/training/scala/site/}

Scala is a programming language for general software applications. Scala has full support for functional programming and a very strong static type system. This allows programs written in Scala to be very concise and thus smaller in size than other general-purpose programming languages. Many of Scala's design decisions were inspired by criticism of the shortcomings of Java.

\section{Installation}

Windows
Step 1. Download scala from http://www.scala-lang.org/downloads

Step 2. Run installer

Step 3. Verify

Open terminal and check which version of scala

$ scala -version

Scala code runner version 2.11.5 -- Copyright 2002-2013, LAMP/EPFL

\section{IDE}

I use IntelliJ IDEA 2016.2 as scala IDE

IntelliJ IDEA Installation Guide
Online IDE
You can use tryscala as an online IDE

http://www.tryscala.com/

\section{Basic Syntax}

Print print
> println("Hello, Scala!");

Hello, Scala!
Conditional
if Statement

if statement consists of a Boolean expression followed by one or more statements.

var x = 10;
if( x < 20 ){
 println("This is if statement");
}
if-else Statement

var x = 30
if( x < 20 ){
  println("This is if statement");
} else {
  println("This is else statement");
}
if-else if-else Statement

 var x = 30;
if( x == 10 ){
 println("Value of X is 10");
} else if( x == 20 ){
 println("Value of X is 20");
} else if( x == 30 ){
 println("Value of X is 30");
} else{
 println("This is else statement");
}
Coding Convention 1
Keep It Simple
Don't pack two much in one expression
/*
 * It's amazing what you can get done in a single statement
 * But that does not mean you have to do it.
 */
jp.getRawClasspath.filter(
  _.getEntryKind == IClasspathEntry.CPE_SOURCE).
  iterator.flatMap(entry =>
    flatten(ResourcesPlugin.getWorkspace.
      getRoot.findMember(entry.getPath)))
Refactor
There's a lot of value in meaningfull names.
Easy to add them using inline vals and defs
val sources = jp.getRawClasspath.filter(
  _.getEntryKind == IClasspathEntry.CPE_SOURCE)
def workspaceRoot =
  ResourcesPlugin.getWorkspace.getRoot
def filesOfEntry(entry: Set[File]) =
  flatten(worspaceRoot.findMember(entry.getPath)
sources.iterator flatMap filesOfEntry
Prefer Functional
By default

use vals, not vars
use recursions or combinators, not loops
use immutable collections
concentrate on transformations, not CRUD
When to deviate from the default - sometimes, mutable gives better performance. - sometimes (but not that often!) it adds convenience

But don't diablolize local state
Why does mutable state lead to complexity?

It interacts with different program parts in ways that are hard to track.

=> Local state is less harmful than global state.

"Var" Shortcuts
var interfaces = parseClassHeader()...
if (isAnnotation) interfaces += ClassFileAnnotation
Refactor

val parsedIfaces = parseClassHeader()
val interfaces =
  if (isAnnotation) parsedIfaces + ClassFileAnnotation
  else parsedIfaces
Martin Odersky - Scala with Style ↩